---
title: "Morphology 3"
subtitle: "Week 10"
author: "Eileen Blum"
date: "Rutgers University </br> March 20, 2018"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["https://www.jvcasillas.com/ru_xaringan/css/rutgers.css", "https://www.jvcasillas.com/ru_xaringan/css/rutgers-fonts.css"]
    nature:
      beforeInit: "https://www.jvcasillas.com/ru_xaringan/js/ru_xaringan.js"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r, load_libraries, message=FALSE, warning=FALSE, echo=FALSE}
library(untidydata)
library(xaringan)
library(plot3D)
library(tidyverse)
```
# Business

- Homework is due **Monday, March 26 at 5:00 pm**
  - all future homeworks will be due on the following Monday rather than Friday

- Reading is Language Files Ch. 4: Morphology

- Final midterm are grades on sakai

---

# Review

- Morphology is the linguistic subfield concerned with words and word formation

- Speakers do not memorize words
  - morphological processes are part of linguistic competence

- Before spring break, we learned some basics about morphological processes

- Let's review!

---
# Review: Morphological Analysis

- How would we break the following words from Michoacan Aztec down into morphemes?

.pull-left[
.pull-left[
nokali    <br>
nokalimes <br>
mokali    
]
.pull-right[
'my house' <br>
'my houses' <br>
'your house'
]]
.pull-right[
.pull-left[
mopelo    <br>
mopelomes <br>
ipelo     
]
.pull-right[
'your dog' <br>
'your dogs' <br>
'his dog'
]] <br>

--

- What is the morpheme for 'house'?

--

  *kali* (Why not *kal* or *kalim*?)

--

- How would you say 'his houses'?

--

  *ikalimes*
  
---

# Morphological types of languages

- It is true that languages use morphology to varying degrees

--

- **Analytic** languages use *mostly* simple words (only roots, no affixes)

For example,|Mandarin|Chinese|(tones|ommitted)|
------------|--------|-------|------|---------|
Example:    |wo      |men    |tan   |qin      |le
Gloss:      |I       |plural |play  |piano    |past <br>
Translation:| 'We played the piano' 

--

- **Synthetic** languages use complex words

--

  - In **fusional** languages, each morpheme carries a lot of information <br>
Ex., Spanish <br>
.pull-left[
.pull-left[
hablo     <br>
hablamos  <br>
habla     <br>
hablé     
]
.pull-right[
'I am speaking' <br>
'We are speaking' <br>
'S/he is speaking' <br>
'I spoke'
]]

--
.pull-right[
.pull-left[
-o    <br>
-amos <br>
-a    <br>
-é    
]
.pull-right[
1st sg present <br>
1st pl present <br>
3rd sg present <br>
1st sg past
]]

---

# Morphological types of languages

- **Synthetic** languages use complex words

  - **Agglutinative** languages continuously add morphemes <br>
Ex., Japanese <br>
.pull-left[
.pull-left[
tabe-ru <br>
tabe-ta <br>
tabe-rare-ru  <br>
tabe-rare-ta  <br>
tabe-rare-mashi-ta  
]
.pull-right[
'to/will eat' <br>
'ate' <br>
'to/will be eaten' <br>
'was eaten' <br>
'was eaten (polite)'
]]

--
.pull-right[
.pull-left[
-ru <br>
-ta past <br>
-rare <br>
-mashi- ]
.pull-right[
non-past <br>
passive <br>
politeness
]] <br>

--

  - In **polysynthetic** languages, nouns are incorporated into verb morphology <br>
Ex., Sora <br>
anin  -ɲam  -jɔ   -te      -n <br>
he    catch fish  non-past do <br>
'He is catching fish'

---

# Morphological types of languages

- You are now familiar with various types of morphemes, morphological processes, and how to discover them

--

- Next, we will practice some morphological analysis


